% INTRODUCTION
%---------------------------
\chapter{Introduction}
This thesis is an exploration of methods used for the measurement of entanglement in large-scale simulations suitable for condensed matter theory.

In this chapter we discuss different measures of entanglement and their scaling.
In Chapter 2 we describe quantum Monte Carlo algorithms in the valence bond basis.
Chapter 3 contains measurements of valence bond entanglement entropy as compared to standard von Neumann entanglement entropy.
In Chapter 4 we develop a method of measuring all Renyi entanglement entropies, excluding the von Neuman entanglement entropy, using valence bond quantum Monte Carlo.
In the final chapter we summarize the results of this research and provide a discussion.

%---------------------------------------------------------------------------------------------------------
\section{Quantum Entanglement}
%---------------------------------------------------------------------------------------------------------
Entanglement is one important and interesting feature that differentiates quantum mechanical systems from classical systems.
In a pair of entangled spins/particles/photons/states the members of the system have information about each other, even if they are spatially separated.
Two spin-1/2 spins, for instance, are considered entangled if their joint state can not be written as the product of their individual states.  Of the following two states,
\begin{equation}
	\ket{\Psi_{1,2}} = \frac{1}{\sqrt{2}} \big( \ket{\up_1\up_2} + \ket{\dw_1\dw_2}\big) 
	\hspace{1cm} \text{and} \hspace{1cm}
	\ket{\Phi_{1,2}} = \frac{1}{\sqrt{2}} \big( \ket{\up_1\dw_2} - \ket{\dw_1\dw_2}\big), 
\end{equation}
only $\ket{\Phi_{1,2}}$ is separable, and can be rewritten as
\begin{equation}
	\ket{\Phi_{1,2}} = \frac{1}{\sqrt{2}}  \big( \ket{\up_1} - \ket{\dw_1}\big) \otimes \ket{\dw_2}.
\end{equation}
Therefore $\ket{\Phi_{1,2}}$ is unentangled, while $\ket{\Psi_{1,2}}$, which cannot be written as a product state, is an entangled state.
In the case of only two spins it is easy to see that states are separable (unentangled). As we add more spins to the system it becomes less apparent which  states are entangled. Furthermore, we can quantify the degree to which they are entangled beyond being either separable states, or non-separable states (entangled).

%---------------------------------------------------------------------------------------------------------
\section{Measures of Entanglement}
%---------------------------------------------------------------------------------------------------------

There are many different quantities that can be used to measure entanglement . Though not exhaustive a few measures include concurrence \cite{concurrence}, logarithmic negativity \cite{negativity}, squashed entanglement \cite{squashed}, and entropy of entanglement \cite{EE}.
Measures of entanglement share three properties \cite{Plenio2005}:
\begin{enumerate}
\item An entanglement measure $E(\rho)$, measuring the entanglement between two spins, is a mapping from density matrices to positive real numbers defined for states of arbitrary bipartite systems.
Usually a normalization factor is included such that the maximally entangled state,
\begin{equation}
\ket{\psi_{\rm max}}  = \frac{\ket{\up\up} + \ket{\dw\dw}}{\sqrt{2}},
\end{equation}
 of two qubits has $E(\ket{\psi_{\rm max}}\bra{\psi_{\rm max}}) = \ln(2)$.
 \item  $E(\rho)=0$ if the state $\rho$ is separable.
 \item $E(\rho)$ does not increase under local operators on either of the spins.  That is, if we have an operator acting on only {\it one} of the spins, the entanglement between the two spins will not increase.
% \item For a pure state $\ket{\psi}\bra{\psi}$ the measure reduces to the von Neumann entanglement entropy.
\end{enumerate}
There is also a fourth condition, not satisfied by all measures of entanglement, requiring the measure to reduce to the von Neumann entanglement entropy for a system in a pure state.


%---------------------------------------------------------------------------------------------------------
\section{The von Neumann Entanglement Entropy}
%---------------------------------------------------------------------------------------------------------
For a system partitioned into two regions, A and B, the von Neumann entanglement entropy \vn is defined as
\begin{equation}
	\VN_\text{A} = -\rm{Tr}\left({\rho_A \ln \rho_A}\right),
\end{equation}
where $\rho_{\rm{A}}=\rm{Tr}_B\ket{\psi}\bra{\psi}$ is the reduced density matrix, the density matrix of the entire system with the degrees of freedom from region B traced out.
\vn is only suited to measure the entanglement of a pure state, as is the case for all the entanglement measures we will discuss in this thesis.
This is due to the way \vn measures entanglement; we begin with a pure state and trace out the degrees of freedom for some subregion B.  
We are left with $\rho_{\rm A}$ representing a (possibly mixed) state.
\vn can be rewritten in terms of the eigenvalues of the density matrix,
\begin{equation}
\VN_\text{A} = -\sum_i \left(\lambda_i \ln \lambda_i\right),
\end{equation}
which is quite similar to the form of thermodynamic entropy or Shannon information entropy. If $\rho_{\rm A}$ represents a pure state, then we {\it know} what state of region A is in, and $\rho_{\rm A}$ will have one non-zero eigenvalue equal to 1 (so $\VN = 0$).
If $\rho_{\rm A}$ represents a mixed state, we will have some probability distribution of the possible states (so $\VN > 0$). The entropy of that distribution is used to quantify the entanglement between regions A and B.
In the maximally entangled case all eigenvalues are equal to $1/d$ (representing an equal probability of encountering each spin state), where $d$ is the dimension of the Hilbert space for region A ($2^N$ for a system of $N$ spin-1/2 particles). Thus the maximum entanglement entropy for a spin-$1/2$ system is $\VN= \ln (d) = N \ln(2)$. 
If the initial state starts off mixed instead of pure, the state would already have some entropy. The entanglement entropy would contain both the initially present classical uncertainty in the state as well as that due to tracing out region B.

The von Neumann entanglement entropy is part of a larger class of entanglement entropies called the generalized Renyi entanglement entropies.
The $n^\text{th}$ Renyi entanglement entropy $S_n$ is defined as
\begin{equation} \label{renyi}
 	S_n(\rho_{\rm A}) = \frac{1}{1-n}\ln\left[{\rm Tr}\left( \rho_{\rm A}^n \right) \right].
\end{equation}
%It may not be immediately apparent that $\VN = S_1$, but if we take the limit as $n\to 1$,
%\begin{equation}
%S_1(\rho_{\rm A}) = \lim_{n\to1}\frac{\ln\left[{\rm Tr}\left( \rho_{\rm A}^n \right) \right]}{1-n}
%	=\lim_{n\to1}\frac{\frac{d}{dn}\ln\left[{\rm Tr}\left( \rho_{\rm A}^n \right) \right]}{\frac{d}{dn}(1-n)}
%\end{equation}
Taking the limit $n\to1$ gives us $S_1 = \VN$.  These entanglement entropies have the property that for $n>m$, $S_n\le S_m$; successive Renyi entropies give a lower bound on the previous entropies \cite{renyiproof}. They are useful measures which fully characterize the entanglement entropy of the system.

%---------------------------------------------------------------------------------------------------------
\section{Scaling of Entanglement Entropy in Condensed Matter Systems}
%---------------------------------------------------------------------------------------------------------

Using measures of entanglement from quantum information to study interacting quantum many-body systems is a rapidly growing interdisciplinary topic \cite{Amico, intro}.
Entanglement has the advantage of measuring ``hidden'' correlations which can occur in states where correlation lengths diverge or decay exponentially.
Additionally, measurement of entanglement is independent of the basis chosen to represent the condensed matter system.


\subsection{Area Law Scaling}

In condensed matter systems, away from special critical points, entanglement entropy typically follows an ``area law'' \cite{Shredder, ALreview}, perhaps more aptly named a ``boundary law'', where the entanglement scales with the size of the boundary between two regions (A and B) of the system.
In non-critical one-dimensional (1D) systems, for example, the entanglement entropy scales as a constant, since the boundary between regions is zero-dimensional.
For two-dimensional (2D) systems we expect the entanglement to scale with the length of the boundary, and in three dimensions the entanglement should scale with the area of the boundary.
If \vn has area law scaling for a given system, all Renyi entanglement entropies will follow the same scaling \cite{Bbob}.
This area law corresponds to systems with mainly short range entanglement, in which only sites ``near'' the boundary contribute to the entanglement between regions.  
If we had a system with extremely long-range entanglement it would be expected that the entire region A is entangled with region B, and one could see a ``volume law'' scaling of entanglement. 
However, this area law is expected to hold for the ground states of many interacting quantum systems.



%---------------------------------------------------------------------------------------------------------
\subsection{1D Critical Systems}
\label{1dcft}
%---------------------------------------------------------------------------------------------------------

The 1D Heisenberg spin chain is critical; in this and other such systems the entanglement entropy is known to scale as \cite{Cardy, Ian1, Zhou2006}
\begin{equation} \label{cft1d} 
	\VN_{\text{PBC}} = \frac{c}{3}\ln(x') + s_1
	\hspace{1cm}
	\VN_{\text{OBC}} = \frac{c}{6}\ln(2x') + \ln(g)+ \frac{s_1}{2}
\end{equation}
for systems with open (OBC) and periodic boundary conditions (PBC), where $c$ is the central charge from conformal field theory, $g$ is a universal boundary term \cite{AffleckAndLudwig}, and $s_1$ is a model dependent constant. $x'$ is the conformal mapping for the number of sites included in region A, used to account for the finite size of the Heisenberg chains, since \eqref{cft1d} applies in the long chain-length limit.  $x \rightarrow x' = (L/\pi)\sin(\pi x/L)$ for PBC and $x \rightarrow 2x'$ for OBC, where $L$ is the length of the chain.
This known scaling is used to verify the accuracy of our entanglement measurements in Chapter 3.

\subsection{Subleading Corrections to the Area Law}
\subsubsection{Topological Entanglement Entropy}

In systems exhibiting topological order -- a type of order not characterized by local order parameters in which the degeneracy of the ground state depends on the topology of the system -- there exists a universal subleading correction to the area law scaling of entanglement entropy called the {\it topological entanglement entropy}.
For 2D systems with topological order, the entanglement entropy is known to scale as  \cite{Bbob, KP}
\begin{equation} \label{AL}
S_{\rm A} = \alpha \ell - n \gamma + \mathcal{O}(1/\ell),
\end{equation}
where $\ell$ is the length of the boundary, $\gamma$ is the topological entanglement entropy, and $n$ is related to the shape of the boundary.
This topological entanglement entropy is zero when the system is not topologically ordered, and  otherwise takes a value related to the degeneracy of the ground state.
Thus it can be thought of as a type of topological order parameter.
By measuring the entanglement entropy of regions of a system with certain geometries which are chosen to cancel out other terms of the entanglement scaling, it is possible to measure the topological entanglement entropy of a system \cite{LW}.

\subsubsection{At Quantum Critical Points}

The scaling of entanglement entropy at quantum critical points is also known to hold information about universal quantities.
In critical systems the area law scaling has universal additive logarithmic corrections \cite{Casini2007,Ryu, max}.
By extracting these corrections, entanglement entropy can be used to to detect and characterize phase transitions.

%\change{New algorithms, such as tensor-network states and projected entangled pair states, rely on ansatz wavefunctions that are explicitly constructed to obey an area law \cite{MERA,PEPS1,PEPS2} but can scale to large system sizes. States scaling no faster than this area law can be accurately represented with matrix product states \cite{MPS_DMRG}. }

\section{Measurement of Entanglement Entropy}

Many methods have been developed that are capable of measuring the entanglement entropy. In 1D density matrix renormalization group (DMRG) simulations and exact diagonalization are particularly powerful tools. They provide direct access to the density matrix and wavefunction. However exact diagonalization scales exponentially poorly as a function of system size and DMRG is primarily a tool useful for 1D and multi-leg ladders. To study and extract these universal quantities in 2D and higher it is necessary to be able to measure entanglement entropy in a scalable type of simulation. 
Quantum Monte Carlo simulations scale well in any dimension and allow a choice of basis. 
%\change{which is not constructed to obey any particular entanglement scaling relationships}.
In this thesis we explore methods of measuring entanglement entropy in quantum Monte Carlo simulations in the valence bond basis.


