% LaTeX Curriculum Vitae Template
%
% Copyright (C) 2004-2009 Jason Blevins <jrblevin@sdf.lonestar.org>
% http://jblevins.org/projects/cv-template/
%
% You may use use this document as a template to create your own CV
% and you may redistribute the source code freely. No attribution is
% required in any resulting documents. I do ask that you please leave
% this notice and the above URL in the source code if you choose to
% redistribute this file.

\documentclass[letterpaper]{article}

\usepackage{hyperref}
\usepackage{geometry}

% Comment the following lines to use the default Computer Modern font
% instead of the Palatino font provided by the mathpazo package.
% Remove the 'osf' bit if you don't like the old style figures.
\usepackage[T1]{fontenc}
\usepackage[sc,osf]{mathpazo}

% Set your name here
\def\name{Ann Berlinsky Kallin}

% Replace this with a link to your CV if you like, or set it empty
% (as in \def\footerlink{}) to remove the link in the footer:
\def\footerlink{http://jblevins.org/projects/cv-template/}

% The following metadata will show up in the PDF properties
\hypersetup{
  colorlinks = true,
  urlcolor = black,
  pdfauthor = {\name},
  pdfkeywords = {economics, statistics, mathematics},
  pdftitle = {\name: Curriculum Vitae},
  pdfsubject = {Curriculum Vitae},
  pdfpagemode = UseNone
}

\geometry{
  body={6.5in, 8.5in},
  left=1.0in,
  top=1.25in
}

% Customize page headers
\pagestyle{myheadings}
\markright{\name}
\thispagestyle{empty}

% Custom section fonts
\usepackage{sectsty}
\sectionfont{\rmfamily\mdseries\Large}
\subsectionfont{\rmfamily\mdseries\itshape\large}

% Other possible font commands include:
% \ttfamily for teletype,
% \sffamily for sans serif,
% \bfseries for bold,
% \scshape for small caps,
% \normalsize, \large, \Large, \LARGE sizes.

% Don't indent paragraphs.
\setlength\parindent{0em}

% Make lists without bullets
\renewenvironment{itemize}{
  \begin{list}{}{
    \setlength{\leftmargin}{1.5em}
  }
}{
  \end{list}
}

\begin{document}

% Place name at left
{\huge \name}

% Alternatively, print name centered and bold:
%\centerline{\huge \bf \name}

\vspace{0.25in}

\begin{minipage}{0.45\linewidth}
  Department of Physics \& Astronomy \\
  \href{http://uwaterloo.ca/}{University of Waterloo} \\
  Waterloo, Ontario, N2L 3G1, Canada
\end{minipage}
\begin{minipage}{0.45\linewidth}
  \begin{tabular}{ll}
    Phone: & (226) 868-0431 \\
%    Fax: &  (919) 962-5678 \\
    Email: & \href{mailto:ann.kallin@gmail.com}{\tt ann.kallin@gmail.com} \\
                & \href{mailto:akallin@uwaterloo.ca}{\tt akallin@uwaterloo.ca} \\
   
%    Homepage: & \href{http://www.stat-or.unc.edu/}{\tt http://www.stat-or.unc.edu/} \\
  \end{tabular}
\end{minipage}


\section*{Personal}

\begin{itemize}
\item Born on September 11, 1985 in Vancouver, BC.
\item Canadian and American citizen.
\end{itemize}


\section*{Education}

\begin{itemize}
	\item Ph.D. in Physics, University of Waterloo, Ontario (In Progress)
  	\item M.Sc. in Physics, University of Waterloo, Ontario (2010)
	\item B.Sc. in Honours Physics, McMaster University, Hamilton, Ontario (2008)
\end{itemize}


\section*{Awards and Honours}

\begin{itemize}

\item Ontario Graduate Scholarship (2013-2014)
\item J. Alan George Student Leadership Award (2011)
\item Alumni Gold Medal Award (2011)
\item NSERC Postgraduate Scholarship (PGS D) (2011-2013)
\item Outstanding Achievement in Graduate Studies Award (Master's) (2010)
\item Ontario Graduate Scholarship (2010) (Declined)
\item CAM Graduate Student Physics Conference Presenter Scholarship (2009)
\item NSERC Postgraduate Scholarship (PGS M) (2008-2010)
\item Ontario Graduate Scholarship (2008) (Declined)
\item NSERC Undergraduate Student Research Award (2008)
\item McMaster University Futures Fund Graduand Award (2008)
\item Harry Lyman Hooker Scholarship (2007)
\item The Boyd McLay Scholarship in Physics (2007)
\item NSERC Undergraduate Student Research Award (2007)
\item Dean's Honour List each undergraduate term, graduated with distinction (2004-2008)
\item The McMaster Honour Entrance Award (2004 - 2005)
\end{itemize}

\section*{International Workshops}

\begin{itemize}

\item {\bf KITP Affiliate} for the program Disentangling Quantum Many-body Systems: Computational and Conceptual Approaches (2010)

\item {\bf Boulder Summer School} on Computational and Conceptual Approaches
to Quantum Many-Body Systems (2010)

\end{itemize}

\section*{Conference Talks and Posters}

\begin{itemize}
\item CECAM Summer School on Quantum Spin Liquids, Trieste, Italy (2013) \\
	Poster --  {\it Measuring Entanglement at a Quantum Critical Point with Numerical Linked Cluster Expansion}
\item CAP Congress, Montreal, Quebec (2013) \\
	Talk --  {\it The Numerical Linked Cluster Expansion Method}
\item APS March Meeting, Baltimore, Maryland (2013) \\
	Talk --  {\it Measuring Entanglement at a Quantum Critical Point with Numerical Linked Cluster Expansion}
\item APS March Meeting, Boston, Massachusetts (2012) \\
	Talk --  {\it Scaling of Entanglement Entropy in the 2D Heisenberg Ground State}
\item Workshop on Quantum Information in Quantum Many-Body Physics, Universit\'e de Montr\'eal  (2011) \\ Talk --  {\it Entanglement Scaling in the 2D Heisenberg Model}
\item KITP program on Disentangling Quantum Many-body Systems, Santa Barbara, California (2010) \\
 	Talk -- {\it Computing Entanglement Scaling in Condensed Matter Ground States}
\item Boulder Summer School for Condensed Matter and Materials Physics, Boulder, Colorado (2010)\\
	Poster -- {\it Measuring Renyi Entanglement Entropy in Quantum Monte Carlo Simulations}
\item APS March Meeting, Portland, Oregon (2010) \\
	Talk --  {\it Measuring Renyi Entanglement Entropy with Quantum Monte Carlo}
\item Waterloo Graduate Student Research Conference, Waterloo, Ontario (2010) \\
	Talk --  {\it Measuring Renyi Entanglement Entropy with Quantum Monte Carlo}
\item Candian-American-Mexican Graduate Student Physics Conference, Acapulco, Mexico (2009) \\
	Talk --  {\it Valence Bond and von Neumann Entanglement Entropy in Quantum Monte Carlo}
\item SHARCNET Research Day, Waterloo, Ontario (2009)  \\
	Poster -- {\it Entanglement Entropy in Quantum Monte Carlo}
\item Canadian Undergraduate Physics Conference, Frederickton, New Brunswick (2006)\\
	Poster -- {\it Crystal Growth: In the Zone}
\item Canadian Undergraduate Physics Conference, London, Ontario (2005)\\
	Poster -- {\it Floating Zone Crystal Growth of High-T$_c$ Superconductors}
\end{itemize}

%\section*{Employment}
%\begin{itemize}
%\item Stanford University 1927--1931.
%\item Columbia University 1931--1946.
%\item University of North Carolina, 1946--1973.
%\end{itemize}


\section*{Publications}

%\subsection*{Journal Articles}

\begin{enumerate}

\item {\it Two Dimensional Incommensurate and Three Dimensional Commensurate Magnetic Order and Fluctuations in $La_{2-x}Ba_{x}CuO_{4}$}\\
J.J. Wagman, G. Van Gastel, K.A. Ross, Z. Yamani, Y. Zhao, Y. Qiu, J.R.D. Copley, A.B. Kallin, E. Mazurek, J. P. Carlo, H.A. Dabkowska, B.D. Gaulin\\
Phys. Rev. B {\bf 88}, 014412 (2013)

\item {\it Detecting Classical Phase Transitions with Renyi Mutual Information}\\
Jason Iaconis, Stephen Inglis, Ann B. Kallin, and Roger G. Melko \\
Phys. Rev. B {\bf 87}, 195134 (2013)

\item {\it Entanglement at a Two-Dimensional Quantum Critical Point: a Numerical Linked Cluster Expansion Study}\\
Ann B. Kallin, Katharine Hyatt, Rajiv R. P. Singh, Roger G. Melko\\
Phys. Rev. Lett. {\bf 110}, 135702 (2013)

\item {\it Entanglement scaling in two-dimensional gapless systems} \\
Hyejin Ju, Ann B. Kallin, Paul Fendley, Matthew B. Hastings, and Roger G. Melko \\
Phys. Rev. B {\bf 85}, 165121 (2012)

\item  {\it Anomalies in the Entanglement Properties of the Square-Lattice Heisenberg Model} \\ %[4 citations] \\
Ann B. Kallin, Matthew B. Hastings, Roger G. Melko, and Rajiv R. P. Singh \\
Phys. Rev. B {\bf 84}, 165134 (2011) 

\item  {\it Finite-Temperature Critical Behavior of Mutual Information}\\% [5 citations]\\
Rajiv R. P. Singh, Matthew B. Hastings, Ann B. Kallin, and Roger G. Melko\\ 
Phys. Rev. Lett. {\bf 106}, 135701 (2011)

\item  {\it Finite-size scaling of mutual information in Monte Carlo simulations: Application to the spin-1/2 XXZ model}\\% [8 citations] \\
Roger G. Melko, Ann B. Kallin, and Matthew B. Hastings  \\
Phys. Rev. B {\bf 82}, 100409(R) (2010) 

\item  {\it Measuring Renyi Entanglement Entropy in Quantum Monte Carlo}\\% [29 citations]\\
Matthew B. Hastings, Ivan Gonzalez, Ann B. Kallin, and Roger G. Melko \\
Phys. Rev. Lett. {\bf 104}, 157201 (2010) 

\item  {\it Valence Bond and von Neumann Entanglement Entropy in Heisenberg Ladders} \\%[23 citations] \\
Ann B. Kallin, Ivan Gonzalez, Matthew B Hastings, and Roger G. Melko\\
Phys. Rev. Lett. {\bf 103}, 117203 (2009)

\item {\it Diagonal and collinear incommensurate spin structures in underdoped La$_{2-x}$Ba$_x$CuO$_4$} \\%[3 citations]\\
    S.R. Dunsiger, Y. Zhao, B.D. Gaulin, Y. Qiu, P. Bourges, Y. Sidis, J.R.D. Copley,  
  A.B. Kallin, E.M. Mazurek, and H.A. Dabkowska \\
Phys. Rev. B {\bf 78}, 092507 (2008) 
\end{enumerate}

\section*{Service}

\begin{itemize}

\item {\bf Outreach} 
	\begin{itemize}
	
	\item Presenter at Wilfrid Laurier LEAP Camp (for grade 7-8 girls) visit to the Perimeter Institute (July 2013)
	\item Instructor for UW DIRECTIONS ESTEEM Conference, an Aboriginal High School Enrichment Program (May 2012)
	\item Volunteer/Instructor at a Canadian Association for Girls in Science (CAGIS) event hosted by the IQC (April 2012)
	\item Volunteer Judge for Centennial Public School Science Fair (March 2012)
	\item Instructor for UW DIRECTIONS ESTEEM Conference (March 2011)
	\end{itemize}

\item {\bf Conference Organizer} for the Waterloo DMRG Winter School, Waterloo (2011)	
\item {\bf Referee} for Physical Review Letters, Physical Review A, Physical Review B, Physica A
\end{itemize}

\section*{Extracurricular Activities}

\begin{itemize}

\item Member of the KW Women's Recreational Hockey League - Team Aardvark (2013 - Present)
\item Captain of the UW Physics intramural ice hockey team - Spin Ice (2012 - Present)
\item Member of Perimeter Institute Monday morning hockey league (2011 - Present)
\item Member of the UW Physics intramural ice hockey team - Spin Ice (2008 - Present)
\item Member of the UW Physics Tough Mudder obstacle course team - The Jason Iaconis Experience (2012)
\item Member of LOCO Roller Derby - Kitchener Chapter (2011)
\item Volunteer for the Tri-City Roller Girls (2011)

\end{itemize}

\bigskip

% Footer
\begin{center}
  \begin{footnotesize}
    Last updated: \today \\
%    \href{\footerlink}{\texttt{\footerlink}}
  \end{footnotesize}
\end{center}

\end{document}