Our paper used large-scale numerical simulations to look at a direct comparison between the von Neumann (VN) entropy (the standard measure of quantum entanglement) and the valence bond (VB)  entropy, which has recently been put forth as an alternative for the VN entropy amenable to QMC simulations in 2D and higher.  Initial results (Ref. [7] and [8]) claimed that many properties of the VB entropy were the same as the VN entropy.  We find however that the valence bond entropy can neither give a bound on the VN entropy, nor can it reproduce universal constants derived from the VN entropy (known from conformal field theory).  We develop an explanation for why the VB entropy is greater than the VN entropy in the  2D Neel groundstate of the Heisenberg model, and demonstrate this fact using our unbiased numerical simulations (QMC and DMRG).

We reproduce previous results showing that VB entropy does *not* obey the area law in the 2D Neel groundstate of the Heisenberg model.  This result, when first reported in the literature (Ref. [7] and [8]), caused intense interest in a broad range of communities, including quantum information, quantum many-body simulation, and condensed-matter physics, since it implies that novel new classes of quantum simulations based on matrix-product states (the so-called PEPS or MERA simulation frameworks) would *not* be able to faithfully represent the prototypical Neel state, because they are constructed to obey the area law.  This has broad and important consequences for any field which desires sign-problem free simulations of 2D (and higher) quantum groundstates (e.g. the Hubbard-model of superconductivity, frustrated quantum spins), a prospect that PEPS and MERA frameworks have perhaps the greatest promise to achieve.

However in our work, we show the first conclusive evidence that the von Neumann entropy *does* obey the area law in this 2D Neel state, contrary to the VB entropy.  Further, we develop an alternate explanation of the valence bond entanglement entropy in terms of the bond length distribution, which we use to demonstrate *why* the valence bond entropy contains a multiplicative log correction in the Neel state.

Therefore, besides providing a firm comparison between the VB and VN entropies plus providing a new explanation for the VB entropy defined via the bond-length distribution, with this work we have fully elucidated the nature of the dramatic correction to the area law in the Neel state defined via the VB entropy.  We have showed that the "true" von Neumann entropy does *not* contain this correction, but obeys the area law. Since our results have far reaching-consequences in the fields of condensed matter, quantum information, and quantum many-body simulations, we feel that PRL is the appropriate venue for its publication.
